Universal Characterization of Variability by \\Shannon's $\sum_ip_i\,\textrm{ln}\,p_i$ Formula

### 2019-09-18 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding entropy as state variable

As we saw in the previous sections, Shannon entropy is a very general concept that applies to any type of distribution, no matter what the problem domain. To understand its link to the thermodynamic/statistical mechanics entropies we need two answer two questions. First, what does the Shannon entropy represents in the context of statistical mechanics? Second, why does the Shannon entropy have the same properties of the thermodynamic entropy? That is, we have to understand why it increases during irreversible evolution, it is conserved during reversible evolution, it is maximized at equilibrium and it is additive for independent systems.

The first question is straight-forward to answer. The macroscopic variables of a system (e.g. volume, pression, energy) do not give a full description of its pars. If we fix a macrostate and then measured the state of all microscopic constituents, we would not find the same answer every time. The choice of a macrostate, then, identify a set of possible microstates that are compatible with it. It is then valid to ask the question: what is the variability of microstates one can find within a given macrostates? This would be answered by the Shannon entropy, which will measure the variability in bits required to identify a microstate given the macrostate.

>>>>

As we saw in the previous sections, Shannon entropy is a very general concept that applies to any type of distribution, no matter what the problem domain. To understand its link to the thermodynamic/statistical mechanics entropies we need two answer two questions. First, what does the Shannon entropy represents in the context of statistical mechanics and why is it a state variable? Second, why does the Shannon entropy have the same properties of the thermodynamic entropy? That is, we have to understand why it increases during irreversible evolution, it is conserved during reversible evolution, it is maximized at equilibrium and it is additive for independent systems.

The first question is straight-forward to answer. The macroscopic variables of a system (e.g. volume, pression, energy) do not give a full description of its pars. If we fix a macrostate and then measured the state of all microscopic constituents, we would not find the same answer every time. The choice of a macrostate, then, identifies a set of possible microstates that are compatible with it. It is then valid to ask the question: what is the variability of microstates one can find within a given macrostates? This would be answered by the Shannon entropy, which will measure the variability in bits required to identify a microstate given the macrostate. This quantity is a property of the macrostate: it is a state variable.
